---
title: "Data Documentation"
author: "Diego A. Dulanto"
date: "3/9/2021"
output: 
  html_document: default
  github_document: default
---


### Data Descriptions

| Variable | Type | Value |
| ----------- | ----------- | ----------- |
| Month of Admission | Numeric | 1-12 |
| Year of Admission | Numeric | 1999-2000 |
| Country of Birth | Numeric | 1-230 |
| Age | Numeric | 0-99 |
| Country of Last Permanent Residence | String | Ex. Peru |
| Marital Status | Numeric | 1-9 |
| State of Intended Residence | Numeric | 1-99 |
| Metropolitan Area of Intended Residence | Numeric | 1-999 |
| Labor Certification | Numeric | 1, 3, 9 |
| Occupation | Numeric | 10-999 |
| Sex | Numeric | 1, 2, 9 |
| Country of Chargeability | String | Ex. Peru |
| Type of Case | Numeric | 1, 3 |
| Nonimmigrant Class of Entry | Numeric | 1-99 |
| Nonimmigrant Year of Entry | Numeric | 1930-2000, 9999|
| Immigrant Class of Admission | Numeric | 1-29 |
| Principal vs. Derivative Class of Admission | Numeric | 1, 2 |

_Missing data will be displayed as NA_

```{r setup}
library(haven)
library(tidyverse)
library(ggplot2)
library(readr)
library(tibble)
```



```{r}
here::here("data", "immigrant_data_2000.sav")

(immigrant_data <- haven::read_sav(
  here::here("data", "immigrant_data_2000.sav")
)
)

##Renaming columns

colnames(immigrant_data)[colnames(immigrant_data) == "V1"] <- "Month of Admission"
colnames(immigrant_data)[colnames(immigrant_data) == "V2"] <- "Year of Admission"
colnames(immigrant_data)[colnames(immigrant_data) == "V3"] <- "Country of Birth"
colnames(immigrant_data)[colnames(immigrant_data) == "V4"] <- "Age"
colnames(immigrant_data)[colnames(immigrant_data) == "V5"] <- "Country of Last Permanent Residence"
colnames(immigrant_data)[colnames(immigrant_data) == "V6"] <- "Marital Status"
colnames(immigrant_data)[colnames(immigrant_data) == "V7"] <- "State of Intended Residence"
colnames(immigrant_data)[colnames(immigrant_data) == "V8"] <- "Metropolitan Area of Intended Residence"
colnames(immigrant_data)[colnames(immigrant_data) == "V9"] <- "Labor Certification"
colnames(immigrant_data)[colnames(immigrant_data) == "V10"] <- "Occupation"
colnames(immigrant_data)[colnames(immigrant_data) == "V11"] <- "Sex"
colnames(immigrant_data)[colnames(immigrant_data) == "V12"] <- "Country of Chargeability"
colnames(immigrant_data)[colnames(immigrant_data) == "V13"] <- "Type of Case"
colnames(immigrant_data)[colnames(immigrant_data) == "V14"] <- "Nonimmigrant class of entry"
colnames(immigrant_data)[colnames(immigrant_data) == "V15"] <- "Nonimmigrant year of entry"
colnames(immigrant_data)[colnames(immigrant_data) == "V16"] <- "Immigrant class of admission"
colnames(immigrant_data)[colnames(immigrant_data) == "V17"] <- "Principal vs. Derivative Class of Admission"


print(immigrant_data)

```



