# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r setup}
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(tidyverse)
library(gridExtra)
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
```

For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?
```{r}
green_data %>% 
  summarise(
    across(
      c(green1:student),
      list(Max = ~ max(.x, na.rm = TRUE),
           Min = ~ min(.x, na.rm = TRUE))))

filter(green_data, id == 5549)

dictionary %>% 
  summarize(dictionary)

green_data %>% 
  ggplot() +
  geom_jitter(aes(x = "id", y = "green"))

#There are -99s in the data that needs to be recoded
#There are duplicate data
```     

  2. Compute total scores for the four scales. 
     Recode variables as needed.
```{r}
temp_recode <- dictionary %>%
  pull(Item)

Q2 <- green_data %>%
  mutate(
    across(
      all_of(temp_recode),
           ~ na_if(.x, -99))) %>%
  mutate(
    id2 = ifelse(
      duplicated(id),
      paste0(id, "r"), id))

reverse_recode <- dictionary %>%
  filter(Keying == -1 | Keying ==-2) %>%
  pull(Item)

Q2.1 <- Q2%>%
  mutate(
    across(
      all_of(reverse_recode),
           ~ recode(.x, 
                    "5" = 1, 
                    "4" = 2, 
                    "3" = 3, 
                    "2" = 4, 
                    "1" = 5)))

Q2.1 %>%
  rowwise() %>%
  group_by(id2) %>%
  mutate(
    Total_green = sum(c_across(green1 : green5),
                      na.rm = TRUE), 
    Total_comp = sum(c_across(comp1 : comp10), 
                     na.rm = TRUE),
    Total_intel = sum(c_across(intel1 : intel10), 
                      na.rm = TRUE),
    Total_open = sum(c_across(open1 : open10), 
                     na.rm = TRUE)) %>% 
  select(id2,
         Total_green,
         Total_comp,
         Total_intel,
         Total_open)

```
     

  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the recaled variables `*_pomp`.
```{r}
Q3 <- Q2.1 %>%
  mutate(
    across(c(green1 : green5),
           ~.x * 4,
           .names = "{.col}_pomp")) %>% 
  mutate(
    across(c(comp1 : open10),
           ~.x * 2, .names = "{.col}_pomp"))

Q3.1 <- Q3 %>%
  rowwise() %>%
  group_by(id2) %>% 
  mutate(Total_green = sum(c_across(green1_pomp : green5_pomp), na.rm = TRUE),
         Total_comp = sum(c_across(comp1_pomp : comp10_pomp), na.rm = TRUE),
         Total_intel = sum(c_across(intel1_pomp : intel10_pomp), na.rm = TRUE),
         Total_open = sum(c_across(open1_pomp : open10_pomp), na.rm = TRUE)) %>% 
  select(id2,
         Total_green,
         Total_comp,
         Total_intel,
         Total_open,
         student)

Q3.1
```
     

  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
```{r}
Q4 <- Q3.1 %>%
  select(id2,
         Total_green, 
         Total_comp, 
         Total_intel, 
         Total_open) %>%
  pivot_longer(., 
               cols = c(Total_green,
                        Total_comp,
                        Total_intel,
                        Total_open), 
               names_to = "POMP_variables", 
               values_to = "POMP_scores")

ggplot(Q4) +
  aes(x = id2, 
      y = POMP_scores) + 
  facet_grid(~ POMP_variables) + 
  geom_point() +
  scale_x_discrete("ID") +
  scale_y_continuous("POMP Scores") + 
  theme_bw()

```
  

  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?
```{r}
Q5 <- Q3.1 %>%
  mutate(
    student = recode(student, 
                     "1" = "Not a student", 
                     "2" = "Student", 
                     .default = "No Response", 
                     .missing = "No Response"))

a <- ggplot(Q5) +
  aes(x = Total_green,
      y = Total_comp, 
      color = student) +
  geom_point() +
  scale_color_manual(values = c("black", 
                                "red", 
                                "blue")) +
  geom_smooth(method = "lm") +
  theme_classic()

b <- ggplot(Q5) +
  aes(x = Total_green,
      y = Total_intel, 
      color = student) +
  geom_point() +
  scale_color_manual(values = c("black", 
                                "red", 
                                "blue")) +
  geom_smooth(method = "lm") +
  theme_classic()

c <- ggplot(Q5) +
  aes(x = Total_green,
      y = Total_open,  
      color = student) +
  geom_point() +
  scale_color_manual(values = c("black", 
                                "red", 
                                "blue")) +
  geom_smooth(method = "lm") +
  theme_classic()

Q5.1 <- Q3.1 %>% 
  mutate(
    student = recode(student, 
                     "1" = "Not a student", 
                     "2" = "Student"))

d <- ggplot(data = subset(Q5.1, 
                          !is.na(student))) +
  aes(x = Total_green,
      y = Total_comp, 
      color = student) +
  scale_color_manual(values = c("red", 
                              "blue")) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic()

e <- ggplot(data = subset(Q5.1, 
                          !is.na(student))) +
  aes(x = Total_green,
      y = Total_intel, 
      color = student) +
  scale_color_manual(values = c("red", 
                              "blue")) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic()

f <- ggplot(data = subset(Q5.1, 
                          !is.na(student))) +
  aes(x = Total_green,
      y = Total_open, 
      color = student) +
  scale_color_manual(values = c("red", 
                              "blue")) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic()

grid.arrange(a,
             b,
             c,
             d,
             e,
             f, 
             nrow = 2)

```
     

  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
```{r}

```
  

  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
```{r}
Q5 %>%
  group_by(student) %>% 
  summarize(across(c(Total_green:Total_open),
                    list(Mu = ~mean(.x,
                                    na.rm = TRUE),
                         Sigma = ~sd(.x,
                                     na.rm=TRUE),
                         Median = ~median(.x,
                                          na.rm=TRUE),
                         Max=~max(.x, 
                                  na.rm = TRUE),
                         Min=~min(.x, 
                                  na.rm = TRUE))))

```
  

  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
