---
title: "hw03"
author: "Diego A. Dulanto"
date: "2/18/2021"
output:
    html_document: default
    github_document: default
    keep_md: TRUE
    theme: paper
---

# Homework 03

```{r setup}
library(dplyr)
library(tidyr)
library(tibble)
library(naniar)
library(scales)
library(ggplot2)
library(gridExtra)
library(tidyverse)
```

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
View(dictionary)
View(green_data)
```


For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?
```{r}
green_data %>% 
  summarise(
    across(
      c(green1:student),
      list(Max = ~ max(.x, na.rm = TRUE),
           Min = ~ min(.x, na.rm = TRUE))))

filter(green_data, id == 5549) #I'm checking for duplicate data

dictionary %>% 
  summarize(dictionary)

green_data %>% 
  ggplot() +
  geom_jitter(aes(x = "id", y = "green"))

#There are -99s in the data that needs to be recoded

recoded_green <-
  green_data %>%
  na_if(-99) %>%
  mutate(
    student = recode(student, "1" = "nonstudent", "2" = "student"))

```     

  2. Compute total scores for the four scales. 
     Recode variables as needed.
```{r}
reversed <- dictionary %>%
  filter(
    Keying == -1 | Keying == -2) %>%
  pull(Item)

reversed

reversed_green <-recoded_green %>%
  mutate(
    across(
      all_of(reversed),
           ~ recode(.x, 
                    "5" = 1, 
                    "4" = 2, 
                    "3" = 3, 
                    "2" = 4, 
                    "1" = 5)))
View(reversed_green)

total_green <-
  reversed_green %>% 
  rowwise() %>% 
  mutate(green_tot = mean(c_across(green1:green5)),
         comp_tot = mean(c_across(comp1:comp10)),
         intel_tot = mean(c_across(intel1:intel10)),
         open_tot = mean(c_across(open1:open10))) %>% 
  na.omit() %>%
  select(id, student, green_tot, comp_tot, intel_tot, open_tot)

View(total_green)

summary(total_green)

```
     

  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the rescaled variables `*_pomp`.
```{r}
rescaled_green <-
  total_green %>%
  mutate(
    green_pomp = rescale(green_tot, 
                         to=c(0,100), 
                         from=c(1,5)),
    comp_pomp = rescale(comp_tot, 
                        to=c(0,100), 
                        from=c(1,5)),
    intel_pomp = rescale(intel_tot, 
                         to=c(0,100), 
                         from=c(1,5)),
    open_pomp = rescale(open_tot, 
                        to=c(0,100), 
                        from=c(1,5))) %>%
  select(id, 
         student, 
         green_pomp, 
         comp_pomp, 
         intel_pomp, 
         open_pomp)

View(rescaled_green)

summary(rescaled_green)

```
     

  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
```{r Q4 fig2, fig.height = 10, fig.width = 15, fig.align = "center"}
green_histo <- ggplot(rescaled_green) +
  aes(x = green_pomp) +
  geom_histogram(fill = "red",
                 bins = 25)+
  theme_classic()

intel_histo  <- ggplot(rescaled_green) +
  aes(x = intel_pomp) +
  geom_histogram(fill = "green",
                 bins = 25)+
  theme_classic()

comp_histo <- ggplot(rescaled_green) +
  aes(x = comp_pomp) +
  geom_histogram(fill = "blue",
                 bins = 25)+
  theme_classic()

open_histo <- ggplot(rescaled_green) +
  aes(x = open_pomp) +
  geom_histogram(fill = "purple",
                 bins = 25)+
  theme_classic()

grid.arrange(green_histo, 
             comp_histo, 
             intel_histo, 
             open_histo)

#The graphs show that these four variables have relatively normal distribution. 
#The data also tells us that, for comp and open, their distribution are slightly skewed left.

```


5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
Include trend lines for **students** and **non-students**.
What do these plots show?
```{r Q5 fig2, fig.height = 10, fig.width = 15, fig.align = "center"}
comp_scatter <-
  ggplot(rescaled_green, aes(x=comp_pomp, 
                            y=green_pomp, 
                            color=student)) + 
  theme_bw() +
  geom_point(alpha = .5) +
  geom_smooth(aes(group=student))

intel_scatter <-
  ggplot(rescaled_green, aes(x=intel_pomp, 
                            y=green_pomp, 
                            color=student)) + 
  theme_bw() +
  geom_point(alpha = .5) +
  geom_smooth(aes(group=student))

open_scatter <-
  ggplot(rescaled_green, aes(x=open_pomp, 
                            y=green_pomp, 
                            color=student)) + 
  theme_bw() +
  geom_point(alpha = .5) +
  geom_smooth(aes(group=student))

grid.arrange(comp_scatter, 
             intel_scatter, 
             open_scatter)

#Intellect and openness have a stronger effect on the green reputation for students.
#The relationship between personality traits and green reputation correlate with students and non-students. 
#The higher the score on compassion, intellect, and openness, the higher the green reputation. 

```

  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
```{r Q6 fig2, fig.height = 10, fig.width = 15, fig.align = "center"}
ggplot(rescaled_green, 
       aes(x = green_pomp, 
           color=student), 
       alpha = .5) +
  geom_bar() +
  geom_density() +
  geom_point(aes(y = "idetity", 
                 color=student), 
             alpha = .5)

#Non-students have a higher count on the green reputation scale than the students. 
#Non-students have a slightly wider distribution on the green reputation scale.

```
  

  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
```{r}
rescaled_green %>%
  select(student, green_pomp, 
         comp_pomp, 
         intel_pomp, 
         open_pomp) %>%
  group_by(student) %>%
  summarize(
    across(c(green_pomp, 
             comp_pomp, 
             intel_pomp, 
             open_pomp),
           list(mean = ~mean(.x, na.rm = TRUE),
                sd = ~sd(.x, na.rm = TRUE),
                med = ~median(.x, na.rm = TRUE),
                min = ~min(.x, na.rm = TRUE),
                max = ~max(.x, na.rm = TRUE))))

#Non-students have a higher green reputation mean, lower standard variance, higher median, and larger range in comparison with students. 
#Notably, non-students have a lower score on openness even though their minimum score is much higher than students. 
#Surprisingly, non-students had a higher minimum score on all personality traits than students.

```
  

  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
